\documentclass[12pt, a4paper]{article}

% --- Packages ---
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{caption}
\usepackage{float}

% --- Colors ---
\definecolor{codeblue}{RGB}{0, 80, 160}
\definecolor{codegray}{RGB}{245, 245, 245}
\definecolor{commentgreen}{RGB}{0, 128, 0}
\definecolor{keywordred}{RGB}{160, 0, 0}

% --- Code style ---
\lstdefinestyle{pythonstyle}{
    backgroundcolor=\color{codegray},
    commentstyle=\color{commentgreen}\itshape,
    keywordstyle=\color{keywordred}\bfseries,
    numberstyle=\tiny\color{gray},
    stringstyle=\color{codeblue},
    basicstyle=\ttfamily\small,
    breaklines=true,
    captionpos=b,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    language=Python,
    frame=single,
    rulecolor=\color{gray!40},
    xleftmargin=1em,
    framexleftmargin=1em,
}
\lstset{style=pythonstyle}

% --- Header / Footer ---
\pagestyle{fancy}
\fancyhf{}
\rhead{Portfolio Risk \& Scenario Engine}
\lhead{Mathematical Reference}
\cfoot{\thepage}

% --- Theorem environments ---
\newtheorem{definition}{Definition}[section]
\newtheorem{remark}{Remark}[section]

% --- Title ---
\title{
    \vspace{-1cm}
    \textbf{\Large Portfolio Risk \& Scenario Engine} \\[0.4em]
    \large Mathematical Documentation \\[0.2em]
    \normalsize A detailed exposition of all computations implemented in the codebase
}
\author{RobinBasso}
\date{\today}

% ============================================================
\begin{document}
% ============================================================

\maketitle
\thispagestyle{fancy}

\begin{abstract}
This document provides a rigorous mathematical description of every computation
performed in the \texttt{portfolio-risk-engine} project. The engine is a
Python-based framework for simulating multi-asset portfolio dynamics and
measuring risk under both normal and stressed market conditions. The four core
modules --- \texttt{portfolio.py}, \texttt{returns\_model.py},
\texttt{monte\_carlo.py}, and \texttt{risk\_metrics.py} --- are covered in full
detail, followed by the scenario framework defined in \texttt{stress\_scenarios.py}.
All formulae are derived from the source code and correspond exactly to the
implemented algorithms.
\end{abstract}

\tableofcontents
\newpage

% ============================================================
\section{Notation and Conventions}
% ============================================================

Throughout this document we adopt the following notation.

\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Symbol} & \textbf{Meaning} \\
\midrule
$N$ & Number of assets in the portfolio \\
$\bm{w} \in \mathbb{R}^N$ & Portfolio weight vector \\
$V_0$ & Initial portfolio value (default: 1\,000\,000\,€) \\
$T$ & Number of time steps (\texttt{n\_steps}) \\
$M$ & Number of simulation paths (\texttt{n\_sims}) \\
$\bm{\mu} \in \mathbb{R}^N$ & Vector of per-step expected returns \\
$\bm{\Sigma} \in \mathbb{R}^{N \times N}$ & Covariance matrix of per-step returns \\
$\bm{r}_t \in \mathbb{R}^N$ & Vector of asset returns at time step $t$ \\
$R_t$ & Scalar portfolio return at time step $t$ \\
$V_t$ & Portfolio value at time step $t$ \\
$\alpha$ & Confidence level for risk metrics (e.g.\ 0.95) \\
\bottomrule
\end{tabular}
\end{center}

All returns in this engine are \textbf{simple (arithmetic) returns}, not
log-returns. Time steps are abstract and dimensionless; they can represent days,
weeks, or any other period depending on how $\bm{\mu}$ and $\bm{\Sigma}$ are
calibrated.

% ============================================================
\section{Portfolio Representation (\texttt{portfolio.py})}
% ============================================================

\subsection{Structure}

A portfolio is fully characterised by the triple $(N, \bm{w}, V_0)$ where
\[
    \bm{w} = (w_1, w_2, \ldots, w_N)^\top, \qquad
    \sum_{i=1}^{N} w_i = 1, \qquad w_i \geq 0 \;\; \forall\, i.
\]
The two constraints --- full investment and no short-selling --- are enforced at
construction time via \texttt{\_validate\_inputs()}.

\subsection{Single-Step Portfolio Return}

Given asset returns $\bm{r}_t = (r_{t,1}, \ldots, r_{t,N})^\top$ at a single
time step $t$, the portfolio return is the weighted inner product:
\begin{equation}
    \boxed{R_t = \bm{w}^\top \bm{r}_t = \sum_{i=1}^{N} w_i \, r_{t,i}}
    \label{eq:portfolio_return}
\end{equation}
This is implemented in \texttt{portfolio\_return} as \texttt{np.dot(weights,
asset\_returns)}.

\subsection{Portfolio Value Path}

Let $\mathbf{R} \in \mathbb{R}^{T \times N}$ be the returns matrix over $T$
steps, where the $t$-th row is $\bm{r}_t^\top$. The scalar portfolio return at
each step is:
\begin{equation}
    R_t = \mathbf{R}[t, :] \cdot \bm{w}, \qquad t = 1, \ldots, T.
\end{equation}
In matrix notation this is simply $\bm{R}_{\text{port}} = \mathbf{R}\,\bm{w}
\in \mathbb{R}^T$.

The portfolio \textbf{value} at time $t$ is computed via the cumulative product
of gross returns, starting from the initial value $V_0$:
\begin{equation}
    \boxed{V_t = V_0 \prod_{s=1}^{t}(1 + R_s)}
    \label{eq:portfolio_path}
\end{equation}
This is the standard compounding formula for arithmetic returns and is
implemented as:
\begin{lstlisting}
portfolio_returns = returns_matrix @ self.weights
return self.initial_value * np.cumprod(1 + portfolio_returns)
\end{lstlisting}
The output array has shape $(T,)$ and contains $[V_1, V_2, \ldots, V_T]$.

\begin{remark}
    The use of \texttt{np.cumprod} over $(1 + R_t)$ is equivalent to the
    discrete-time wealth process. For small returns $R_t \ll 1$, this
    approximates geometric Brownian motion. There is no log-transform applied;
    the model operates in the linear return space.
\end{remark}

% ============================================================
\section{Multivariate Returns Model (\texttt{returns\_model.py})}
% ============================================================

\subsection{Model Specification}

The engine models asset returns using a \textbf{multivariate Gaussian (normal)
distribution}:
\begin{equation}
    \boxed{\bm{r}_t \sim \mathcal{N}(\bm{\mu},\, \bm{\Sigma})}
    \label{eq:returns_model}
\end{equation}
where the same distribution is applied i.i.d.\ at every time step. This implies:
\begin{itemize}[noitemsep]
    \item Returns are serially uncorrelated.
    \item The marginal return of asset $i$ has mean $\mu_i$ and variance $\Sigma_{ii}$.
    \item The co-movement between assets $i$ and $j$ is captured by $\Sigma_{ij}$.
\end{itemize}

\subsection{Covariance Matrix Validation}

The constructor enforces two structural conditions on $\bm{\Sigma}$:
\begin{enumerate}[noitemsep]
    \item \textbf{Squareness:} $\bm{\Sigma} \in \mathbb{R}^{N \times N}$.
    \item \textbf{Symmetry:} $\bm{\Sigma} = \bm{\Sigma}^\top$ (checked via
          \texttt{np.allclose}).
\end{enumerate}
Positive semi-definiteness is not explicitly checked in the code but is
implicitly required for \texttt{np.random.multivariate\_normal} to produce
valid samples.

\subsection{Simulation Algorithm}

For $M$ paths and $T$ steps, the simulation draws:
\begin{equation}
    \bm{r}^{(m)}_t \overset{\text{i.i.d.}}{\sim} \mathcal{N}(\bm{\mu},\,\bm{\Sigma}),
    \qquad m = 1,\ldots,M, \quad t = 1,\ldots,T.
\end{equation}
The output has shape $(M, T, N)$: $M$ simulations, each containing $T$ rows of
$N$-dimensional return vectors. This is a single call to NumPy's
\texttt{multivariate\_normal} with \texttt{size=(n\_sims, n\_steps)}, which
internally uses the Cholesky decomposition $\bm{\Sigma} = \bm{L}\bm{L}^\top$
and transforms standard normal draws:
\begin{equation}
    \bm{r}_t = \bm{\mu} + \bm{L}\, \bm{z}_t, \qquad \bm{z}_t \sim \mathcal{N}(\bm{0}, \bm{I}_N).
\end{equation}

\subsection{Default Calibration (from tests and dashboard)}

The project uses the following default parameters throughout:
\[
    \bm{\mu} = \begin{pmatrix} 0.0005 \\ 0.0002 \end{pmatrix}, \qquad
    \bm{\Sigma} = \begin{pmatrix} 0.0001 & 0.00002 \\ 0.00002 & 0.00008 \end{pmatrix}.
\]
These correspond to a two-asset portfolio (Equity and Bonds). The implied
per-step standard deviations are $\sigma_{\text{Eq}} = 0.01$ (1\%) and
$\sigma_{\text{Bd}} = 0.00894$ (0.894\%), with a correlation of:
\begin{equation}
    \rho_{\text{Eq,Bd}} = \frac{\Sigma_{12}}{\sqrt{\Sigma_{11}\,\Sigma_{22}}}
    = \frac{0.00002}{\sqrt{0.0001 \times 0.00008}}
    = \frac{0.00002}{0.00002\sqrt{2}}
    = \frac{1}{\sqrt{2}} \approx 0.707.
\end{equation}

% ============================================================
\section{Monte Carlo Engine (\texttt{monte\_carlo.py})}
% ============================================================

\subsection{Overview}

The \texttt{MonteCarloEngine} orchestrates the full simulation pipeline. It
combines the \texttt{ReturnsModel} and \texttt{Portfolio} objects to produce an
ensemble of portfolio value paths.

\subsection{Path Simulation}

For each simulation $m \in \{1, \ldots, M\}$:
\begin{enumerate}
    \item Draw an $(T \times N)$ matrix of asset returns
          $\mathbf{R}^{(m)}$ from \texttt{returns\_model.simulate()}.
    \item (Optionally) apply a scenario transformation: $\mathbf{R}^{(m)}
          \leftarrow \mathcal{S}(\mathbf{R}^{(m)})$ (see Section~\ref{sec:scenarios}).
    \item Compute the portfolio value path using equation~\eqref{eq:portfolio_path}:
          \[
            V^{(m)}_t = V_0 \prod_{s=1}^{t}(1 + \bm{w}^\top\bm{r}^{(m)}_s),
            \qquad t = 1,\ldots,T.
          \]
\end{enumerate}

The full output is the matrix $\mathbf{V} \in \mathbb{R}^{M \times T}$, where
entry $(m, t)$ is the portfolio value of path $m$ at time step $t$.

\subsection{Final Value Distribution}

The method \texttt{final\_values} extracts the terminal column:
\begin{equation}
    \boxed{\widetilde{V}^{(m)} = V^{(m)}_T = V_0 \prod_{t=1}^{T}(1 + R^{(m)}_t),
    \qquad m = 1,\ldots,M.}
\end{equation}
This yields a sample $\{\widetilde{V}^{(1)}, \ldots, \widetilde{V}^{(M)}\}$
from the distribution of terminal portfolio wealth, which is then used to
compute risk metrics.

\begin{remark}
    Reproducibility is guaranteed via \texttt{np.random.seed(seed)} before
    sampling. The default seed used throughout the test suite is 42.
\end{remark}

% ============================================================
\section{Risk Metrics (\texttt{risk\_metrics.py})}
% ============================================================

All risk metrics are computed from the empirical distribution of terminal
portfolio values $\{\widetilde{V}^{(m)}\}_{m=1}^{M}$.

\subsection{Value at Risk (VaR)}

\begin{definition}[Value at Risk]
    For a confidence level $\alpha \in (0, 1)$, the \textbf{Value at Risk}
    $\mathrm{VaR}_\alpha$ is the minimum loss not exceeded with probability
    $\alpha$. Equivalently, it is the loss exceeded with probability $1-\alpha$.
\end{definition}

The implementation approximates the initial value by the empirical mean of the
terminal distribution:
\begin{equation}
    \hat{V}_0 = \frac{1}{M}\sum_{m=1}^{M}\widetilde{V}^{(m)}.
\end{equation}
The VaR is then computed as the gap between this mean and the
$(1-\alpha)$-quantile of the terminal distribution:
\begin{equation}
    \boxed{\mathrm{VaR}_\alpha = \hat{V}_0 - q_{1-\alpha}\!\left(\{\widetilde{V}^{(m)}\}\right)}
    \label{eq:var}
\end{equation}
where $q_p(\cdot)$ denotes the $p$-th percentile. For $\alpha = 0.95$, this
extracts the 5th percentile of terminal values:
\[
    \mathrm{VaR}_{0.95} = \hat{V}_0 - q_{0.05}\!\left(\{\widetilde{V}^{(m)}\}\right).
\]
Implemented as:
\begin{lstlisting}
initial_value = np.mean(final_values)
percentile = np.percentile(final_values, (1 - alpha) * 100)
return initial_value - percentile
\end{lstlisting}

\begin{remark}
    Using $\hat{V}_0 = \mathbb{E}[\widetilde{V}]$ rather than the true initial
    value $V_0$ is an approximation. For positively drifted processes ($\mu > 0$),
    $\hat{V}_0 > V_0$, which slightly inflates the reported VaR. A more standard
    convention would use $V_0$ directly.
\end{remark}

\subsection{Conditional Value at Risk (CVaR / Expected Shortfall)}

\begin{definition}[Conditional Value at Risk]
    The \textbf{CVaR} at level $\alpha$ (also called \emph{Expected Shortfall})
    is the expected loss conditional on the loss exceeding $\mathrm{VaR}_\alpha$:
    \[
        \mathrm{CVaR}_\alpha = \mathbb{E}\!\left[\text{Loss} \mid \text{Loss} \geq \mathrm{VaR}_\alpha\right].
    \]
\end{definition}

Let $q^* = q_{1-\alpha}(\{\widetilde{V}^{(m)}\})$ be the VaR threshold. The
tail set is:
\[
    \mathcal{T} = \left\{m : \widetilde{V}^{(m)} \leq q^*\right\}.
\]
The CVaR is computed as the mean loss in this tail:
\begin{equation}
    \boxed{\mathrm{CVaR}_\alpha = \frac{1}{|\mathcal{T}|}\sum_{m \in \mathcal{T}}
    \left(\hat{V}_0 - \widetilde{V}^{(m)}\right)}
    \label{eq:cvar}
\end{equation}
Implemented as:
\begin{lstlisting}
var_threshold = np.percentile(final_values, (1 - alpha) * 100)
tail_losses   = final_values[final_values <= var_threshold]
return np.mean(np.mean(final_values) - tail_losses)
\end{lstlisting}

CVaR is a \textbf{coherent risk measure} (Artzner et al., 1999), satisfying
sub-additivity, monotonicity, positive homogeneity, and translation invariance.
It is always at least as large as $\mathrm{VaR}_\alpha$ by construction:
$\mathrm{CVaR}_\alpha \geq \mathrm{VaR}_\alpha$.

\subsection{Maximum Drawdown}

\begin{definition}[Maximum Drawdown]
    Given a portfolio value path $(V_1, V_2, \ldots, V_T)$, the
    \textbf{maximum drawdown} is:
    \[
        \mathrm{MDD} = \max_{t \in \{1,\ldots,T\}} \frac{\max_{s \leq t} V_s - V_t}
                    {\max_{s \leq t} V_s}.
    \]
\end{definition}

This measures the largest peak-to-trough decline as a proportion of the running
peak. The implementation computes:
\begin{equation}
    \boxed{\mathrm{MDD} = \max_{t}\,\frac{\mathrm{Peak}_t - V_t}{\mathrm{Peak}_t}}
    \label{eq:mdd}
\end{equation}
where $\mathrm{Peak}_t = \max_{s \leq t} V_s$ is the running maximum, obtained
via \texttt{np.maximum.accumulate(path)}:
\begin{lstlisting}
peak      = np.maximum.accumulate(path)
drawdowns = (peak - path) / peak
return np.max(drawdowns)
\end{lstlisting}

\begin{remark}
    MDD is computed on a \emph{single} simulation path (shape $(T,)$), not on
    the terminal distribution. In \texttt{test\_risk\_metrics.py}, it is applied
    to the first path of a 1-simulation run.
\end{remark}

% ============================================================
\section{Stress Testing Framework (\texttt{scenarios/})}
% ============================================================
\label{sec:scenarios}

\subsection{Design Pattern}

The scenario framework uses a \textbf{Strategy design pattern}. The abstract
base class \texttt{Scenario} (in \texttt{base.py}) defines a single interface
method:
\[
    \mathcal{S}: \mathbb{R}^{T \times N} \to \mathbb{R}^{T \times N},
    \qquad \mathbf{R} \mapsto \mathcal{S}(\mathbf{R}).
\]
Concrete scenarios override \texttt{apply(returns)} to perturb the simulated
return matrix before it is passed to the portfolio compounding step.

\subsection{Equity Crash Scenario (\texttt{EquityCrashScenario})}

An instantaneous additive shock is applied to the equity asset (asset index 0)
at the first time step only:
\begin{equation}
    \boxed{\widetilde{r}^{(m)}_{1,0} = r^{(m)}_{1,0} + \delta, \qquad \delta = -0.2}
    \label{eq:equity_crash}
\end{equation}
with all other steps and assets unchanged:
\[
    \widetilde{r}^{(m)}_{t,i} = r^{(m)}_{t,i} \quad \text{for } (t,i) \neq (1,0).
\]
The parameter $\delta = -0.2$ corresponds to an instantaneous \textbf{20\% crash}
in the equity position at $t=1$. In code:
\begin{lstlisting}
shocked[0, 0] += self.crash_size   # crash_size = -0.2 by default
\end{lstlisting}
The impact on portfolio value after one step is:
\[
    V^{\text{stressed}}_1 = V_0\bigl(1 + w_{\text{Eq}}(\mu_{\text{Eq}} + \delta)
    + w_{\text{Bd}}\,r_{1,\text{Bd}}\bigr).
\]
With the default weights $(w_{\text{Eq}}, w_{\text{Bd}}) = (0.6, 0.4)$ and
$\delta = -0.2$, the first-step equity contribution drops by
$0.6 \times 0.2 = 0.12$, i.e.\ a 12\% instantaneous portfolio loss.

\subsection{Volatility Spike Scenario (\texttt{VolatilitySpikeScenario})}

All returns across all time steps and all assets are scaled by a constant factor
$\lambda > 1$:
\begin{equation}
    \boxed{\widetilde{\mathbf{R}}^{(m)} = \lambda \cdot \mathbf{R}^{(m)}}
    \label{eq:vol_spike}
\end{equation}
The default factor is $\lambda = 2$. This multiplicative scaling has the
following effect on the moments of the portfolio return distribution:
\begin{align}
    \mathbb{E}[\widetilde{R}_t] &= \lambda\,\mathbb{E}[R_t] = \lambda\,\bm{w}^\top\bm{\mu}, \\
    \mathrm{Var}[\widetilde{R}_t] &= \lambda^2\,\mathrm{Var}[R_t]
        = \lambda^2\,\bm{w}^\top\bm{\Sigma}\bm{w}.
\end{align}
Thus volatility (standard deviation) is multiplied by $\lambda$, while expected
return is also scaled. This is a \textit{symmetric} stress: both upside and
downside paths are amplified equally.

\subsection{Combined Stress Scenario (\texttt{CombinedStressScenario})}

This scenario applies both an equity crash and a volatility spike
sequentially:
\begin{equation}
    \widetilde{\mathbf{R}}^{(m)} = \lambda \cdot \mathcal{S}_{\text{crash}}(\mathbf{R}^{(m)}),
    \qquad \lambda = 1.5,\;\; \delta = -0.2.
\end{equation}
Explicitly:
\begin{equation}
    \boxed{\widetilde{r}^{(m)}_{t,i} =
    \begin{cases}
        1.5\,(r^{(m)}_{1,0} - 0.2) & \text{if } t = 1,\, i = 0 \\
        1.5\, r^{(m)}_{t,i}         & \text{otherwise}
    \end{cases}}
    \label{eq:combined_stress}
\end{equation}
In code:
\begin{lstlisting}
shocked[0, 0] -= 0.2   # equity crash at t=0
shocked *= 1.5          # scale all returns
\end{lstlisting}

\begin{remark}
    The ordering matters: the crash shock is applied \emph{before} the scaling,
    so the crash magnitude itself is also amplified. The effective first-step
    equity shock becomes $1.5 \times (-0.2) = -0.30$, i.e.\ a 30\% crash on
    the equity position at $t=1$.
\end{remark}

% ============================================================
\section{End-to-End Computational Pipeline}
% ============================================================

The full computation performed for a single simulation run can be summarised
as follows. Let $M$ be the number of paths, $T$ the number of steps, $N$ the
number of assets.

\bigskip
\noindent\textbf{Step 1 — Asset return sampling.}
\[
    \mathbf{R}^{(m)} \sim \mathcal{N}_{T \times N}(\bm{1}_T \bm{\mu}^\top,\,\bm{\Sigma}),
    \quad m = 1, \ldots, M.
\]

\noindent\textbf{Step 2 — Scenario perturbation (optional).}
\[
    \mathbf{R}^{(m)} \leftarrow \mathcal{S}(\mathbf{R}^{(m)}).
\]

\noindent\textbf{Step 3 — Scalar portfolio returns.}
\[
    \bm{R}^{(m)}_{\text{port}} = \mathbf{R}^{(m)} \bm{w} \in \mathbb{R}^T.
\]

\noindent\textbf{Step 4 — Portfolio value paths.}
\[
    V^{(m)}_t = V_0 \prod_{s=1}^{t}(1 + R^{(m)}_{\text{port},s}), \quad t = 1,\ldots,T.
\]

\noindent\textbf{Step 5 — Terminal value extraction.}
\[
    \widetilde{V}^{(m)} = V^{(m)}_T, \quad m = 1,\ldots,M.
\]

\noindent\textbf{Step 6 — Risk metrics computation.}
\begin{align*}
    \hat{V}_0 &= \frac{1}{M}\sum_{m=1}^{M}\widetilde{V}^{(m)}, \\
    \mathrm{VaR}_\alpha &= \hat{V}_0 - q_{1-\alpha}(\{\widetilde{V}^{(m)}\}), \\
    \mathrm{CVaR}_\alpha &= \hat{V}_0 - \frac{1}{|\mathcal{T}|}\sum_{m \in \mathcal{T}}\widetilde{V}^{(m)},
    \quad \mathcal{T} = \{m : \widetilde{V}^{(m)} \leq q_{1-\alpha}\}.
\end{align*}

% ============================================================
\section{Portfolio Return Distribution: Analytical Properties}
% ============================================================

Under the Gaussian returns model, the single-step portfolio return $R_t =
\bm{w}^\top\bm{r}_t$ is itself Gaussian:
\begin{equation}
    R_t \sim \mathcal{N}\!\left(\mu_p,\, \sigma_p^2\right),
\end{equation}
where:
\begin{equation}
    \mu_p = \bm{w}^\top\bm{\mu}, \qquad \sigma_p^2 = \bm{w}^\top\bm{\Sigma}\bm{w}.
    \label{eq:port_moments}
\end{equation}

For the default parameterisation with $\bm{w} = (0.6, 0.4)^\top$:
\begin{align}
    \mu_p &= 0.6 \times 0.0005 + 0.4 \times 0.0002 = 0.000380, \\
    \sigma_p^2 &= (0.6)^2 \times 0.0001
              + 2 \times 0.6 \times 0.4 \times 0.00002
              + (0.4)^2 \times 0.00008 \notag \\
              &= 0.000036 + 0.0000096 + 0.0000128 = 0.0000584, \\
    \sigma_p &\approx 0.00764.
\end{align}

The terminal portfolio value $V_T = V_0\prod_{t=1}^T(1+R_t)$ is the product of
$T$ i.i.d.\ random variables. For large $T$ and small returns, this
approximates a \textbf{log-normal distribution} by the central limit theorem
applied to log-returns:
\[
    \log V_T \approx \log V_0 + T\!\left(\mu_p - \tfrac{1}{2}\sigma_p^2\right)
    + \sigma_p\sqrt{T}\,Z, \quad Z \sim \mathcal{N}(0,1).
\]
The empirical histogram plotted in the Streamlit dashboard reflects this
approximately log-normal shape.

% ============================================================
\section{Module Architecture Summary}
% ============================================================

\begin{table}[H]
\centering
\caption{Summary of modules and their mathematical roles}
\begin{tabular}{llp{8cm}}
\toprule
\textbf{Module} & \textbf{Class / Method} & \textbf{Mathematical Operation} \\
\midrule
\texttt{portfolio.py} & \texttt{portfolio\_return} & $R_t = \bm{w}^\top\bm{r}_t$ \\
& \texttt{portfolio\_path} & $V_t = V_0\prod_{s\leq t}(1+R_s)$ \\
\midrule
\texttt{returns\_model.py} & \texttt{simulate} & $\bm{r}_t \sim \mathcal{N}(\bm{\mu}, \bm{\Sigma})$ \\
\midrule
\texttt{monte\_carlo.py} & \texttt{simulate} & Full path ensemble $\mathbf{V} \in \mathbb{R}^{M\times T}$ \\
& \texttt{final\_values} & Terminal slice $\widetilde{V}^{(m)} = V^{(m)}_T$ \\
\midrule
\texttt{risk\_metrics.py} & \texttt{var} & $\hat{V}_0 - q_{1-\alpha}$ \\
& \texttt{cvar} & Mean loss in tail $\mathcal{T}$ \\
& \texttt{max\_drawdown} & $\max_t (\mathrm{Peak}_t - V_t)/\mathrm{Peak}_t$ \\
\midrule
\texttt{stress\_scenarios.py} & \texttt{EquityCrashScenario} & $r_{1,0} \mathrel{+}= \delta$ \\
& \texttt{VolatilitySpikeScenario} & $\mathbf{R} \leftarrow \lambda \mathbf{R}$ \\
& \texttt{CombinedStressScenario} & Crash then scale \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================
\section{Limitations and Potential Extensions}
% ============================================================

The current implementation makes several simplifying assumptions worth noting:

\begin{enumerate}
    \item \textbf{Gaussian returns.} The model does not capture fat tails,
          skewness, or volatility clustering (e.g.\ GARCH effects). Real asset
          returns exhibit kurtosis significantly greater than 3.

    \item \textbf{Fixed weights.} The portfolio is assumed to be statically
          allocated with no rebalancing. Dynamic strategies (momentum,
          mean-reversion) are not modelled.

    \item \textbf{I.i.d.\ steps.} Returns are independent across time steps.
          Serial correlation and regime-switching are absent.

    \item \textbf{VaR approximation.} Using $\hat{V}_0 = \mathbb{E}[\widetilde{V}]$
          rather than the true $V_0$ introduces a small upward bias in all risk
          metrics.

    \item \textbf{Scenario scope.} Stress scenarios modify only the simulated
          return matrix rather than recalibrating $\bm{\mu}$ and $\bm{\Sigma}$,
          which is a reduced-form rather than structural approach.

    \item \textbf{No discounting.} All values are nominal; there is no risk-free
          rate or discounting applied to future cash flows.
\end{enumerate}

Natural extensions would include: Student-$t$ copulas for fat-tailed dependence,
GARCH(1,1) volatility dynamics, Cholesky-based correlation shocks for
multi-factor stress scenarios, and CVaR-based portfolio optimisation.

% ============================================================
\section*{References}
% ============================================================

\begin{itemize}[noitemsep]
    \item Artzner, P., Delbaen, F., Eber, J.-M., \& Heath, D. (1999).
          \textit{Coherent measures of risk}. Mathematical Finance, 9(3), 203--228.
    \item McNeil, A.J., Frey, R., \& Embrechts, P. (2015).
          \textit{Quantitative Risk Management: Concepts, Techniques and Tools}
          (Revised ed.). Princeton University Press.
    \item Glasserman, P. (2004).
          \textit{Monte Carlo Methods in Financial Engineering}. Springer.
    \item Rannacher, R. (1984). Finite element solution of diffusion problems
          with irregular data. \textit{Numerische Mathematik}, 43, 309--327.
\end{itemize}

\end{document}
